global-model (ODE)

### 2015-02-23 documentation/interactive_spexmodel_attractor.Rmd
interactive simulation of attractor (Shiny)

---
title: "global-model (ODE)"
author: "Florian Schneider"
date: "Thursday, January 30, 2015"
output: html_document
runtime: shiny
---

@@

@@

### initial conditions
```{r, echo=FALSE}

inputPanel(
  selectInput("gradient", label = "select gradient:",             choices = c("b", "L"), selected = "b")
  )

@@

 if(input$gradient == "b") parms$b <- seq(0,1,.005) 
 if(input$gradient == "L") parms$L <- seq(0,20,.1)

parms$rho_ini <- c(0.9, 0.01)

iterations <- expand.grid(parms)
iterations <- cbind(ID = 1:dim(iterations)[1],iterations)

iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]

foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")

  model_parms <- as.list(iterations[iteration,])

  # running the ode-solver
  runmodel <- ode(y = ini_rho(model_parms$rho_ini), func = odesys_spex, times = c(1,1000), parms = model_parms)

  # transfer into ouput and calculate missing rho values
  return(as.data.frame(runmodel)[dim(runmodel)[1],])
  } -> output

output <- cbind(output, iterations)

 if(input$gradient == "b") plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
 if(input$gradient == "L") plot(rho_1 ~ L, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))

>>>>

---
title: "global-model (ODE)"
author: "Florian Schneider"
date: "Thursday, January 30, 2015"
output:
  html_document:
    fig_height: 6.5
runtime: shiny
---

@@

defplot <- function(...) plot(..., 
            type = "l", 
            bty = "l", cex = 0.7, las = 1,
            xlim = c(0,1), ylim = c(0,0.25), 
            xaxs = "i", yaxs = "i", 
            xaxp = c(0,1,5), yaxp = c(0,0.25,5), 
            xlab = "vegetation cover")

rho <- seq(0,1,length = 100)

defaultparms <- list(
    m = 0.05,  # intrinsic mortality of plants 
    r = 1,   # growth rate
    f = 0.8, #  local facilitation
    b = 0.25,  # environmental quality inverse of aridity
    K = 0.9,  # carrying capacity
    alpha = 0.1, # water runoff
    c = 0.2, # local competition
    a = 0.6, # search efficiency
    v = 0.1, # attraction effect
    h = 50, # handling time (one individual on landscape unit)
    p = 0.8, # protection effect
    L = 6, # livestock units per landscape unit 
    q = 1 # search ineffiency at low cover
  )

ini_rho <- function(rho_1, rho_11 = NULL) {
    if(is.null(rho_11)) out <- {
      c(
    rho_1 = rho_1,
    rho_11 = rho_1*rho_1,
    rho_10 = rho_1*(1-rho_1)*2,
    rho_00 = (1-rho_1)*(1-rho_1),
    rho_0 = 1-rho_1
    ) 
    } else {
      out <- c(
    rho_1 = rho_1,
    rho_11 = rho_11,
    rho_10 = rho_1-rho_11,
    rho_00 = 1-2*rho_1+rho_11,
    rho_0 = 1-rho_1
    )
    if(any(out < 0)) out <- c(
    rho_1 = NA,
    rho_11 = NA,
    rho_10 = NA,
    rho_00 = NA,
    rho_0 = NA
    )
    }

  return(out)
}

G <- function(rho_1, q_01 = "auto", prms ) {
  if(q_01 == "auto") q_01 <- 1
  with(prms, r*(b + (1-b)*f*q_01)*rho_1^(1+alpha)*(1-rho_1/(K * (1-c * q_01) ) ))

C <- function(rho_1, q_11 = 1, prms ) with(prms, (m * rho_1) +( (a+v*q_11  ) *rho_1^(1+q)*L*(1-p*q_11))/(1+(a+v*q_11)*h*(rho_1^(1+q)) ) )

attractor <- function(parms, rho_1_ini = seq(0,1, length = 41), rho_11_ini = seq(0,1, length = 11)) {

ini <- list(
  rho_1 = rho_1_ini,
  rho_11 = rho_11_ini,
  rho_10 = NA,
  rho_00 = NA,
  rho_0 = NA
  )

ini <- expand.grid(ini)

for(x in 1:nrow(ini)) {
      temp <- ini_rho(ini[x,]$rho_1, ini[x,]$rho_11)
      ini[x,]$rho_1 <- temp[1]
      ini[x,]$rho_11 <- temp[2]
      ini[x,]$rho_10 <- temp[3]
      ini[x,]$rho_00 <- temp[4]
      ini[x,]$rho_0 <- temp[5]
      }

ini$m_ini <- C(ini$rho_1, ini$rho_11/ini$rho_1, parms)
ini$g_ini <- G(ini$rho_1, (ini$rho_1-ini$rho_11)/(1-ini$rho_1), parms)

ini <- subset(ini, !is.na(rho_1))
ini <- cbind(ID = 1:nrow(ini),ini)

foreach(iteration = ini$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
  source("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock/code/simfunctions.r")

   rho_starting <- ini[iteration, 2:6]

  # running the ode-solver
  runmodel <- ode(y = as.numeric(rho_starting), func = odesys_spex, times = seq(1,2), parms = parms)

  out <-as.data.frame(runmodel)[dim(runmodel)[1],]

  # transfer into ouput and calculate missing rho values
  return(out)
  } -> output

names(output) <- c("time", "rho_1", "rho_11", "rho_10", "rho_00", "rho_0")

rho <- seq(0,1,length = 100)

defplot(rho, C(rho, 1, parms), ylab = "mortality", col= "#00000020")
lines(rho,C(rho, 0.75, parms), col= "#00000040")
lines(rho,C(rho, 0.5, parms), col= "#00000080")
lines(rho,C(rho, 0.25, parms), col= "#00000040")
lines(rho,C(rho, 0, parms), col= "#00000020")

lines(rho, G(rho, 1, parms), col = "#00993320")
lines(rho,G(rho, 0.75, parms), col = "#00993340")
lines(rho,G(rho, 0.5, parms), col = "#00993380")
lines(rho,G(rho, 0.25, parms), col = "#00993340")
lines(rho,G(rho, 0, parms), col = "#00993320")

arrows(ini$rho_1, C(ini$rho_1, ini$rho_11/ini$rho_1, parms),output$rho_1, C(output$rho_1, output$rho_11/output$rho_1, parms), length = 0.04)

arrows(ini$rho_1, G(ini$rho_1, (ini$rho_1-ini$rho_11)/(1-ini$rho_1), parms),output$rho_1, G(output$rho_1, (output$rho_1-output$rho_11)/(1-output$rho_1), parms ), length = 0.02, col = "#009933" )

high_equ <- as.data.frame(ode(y = ini_rho(0.9), func = odesys_spex, times = c(1,1000), parms = parms) )[2,]

points(high_equ$rho_1, C(high_equ$rho_1, high_equ$rho_11/high_equ$rho_1, parms), cex = 2, pch = 20, col = 2)

low_equ <- as.data.frame(ode(y = ini_rho(0.0001), func = odesys_spex, times = c(1,1000), parms = parms) )[2,]

points(low_equ$rho_1, C(low_equ$rho_1,low_equ$rho_11/low_equ$rho_1, parms), pch = 20, cex = 2, col = 2)

@@

### initial conditions
```{r, echo=FALSE, fig.height=7.5}

@@

# if(input$gradient == "b") parms$b <- seq(0,1,.005) 
# if(input$gradient == "L") parms$L <- seq(0,20,.1)

#iterations$f[iterations$f < iterations$b] <- iterations$b[iterations$f < iterations$b]

attractor(parms)

# if(input$gradient == "b") plot(rho_1 ~ b, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
# if(input$gradient == "L") plot(rho_1 ~ L, data = cbind(output, iterations), type = "p", pch  = 20, ylim = c(0,1))
